Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.Rmd
mention the regressor dist is assumed to be the same

(use uniform effect size? normal effect size? -> other dist?)

(average out dist for effect size calc?)

@@

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
\end{equation}
```
\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. The effect size is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.

@@

\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any variables leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients. A conditional expectation is taken because $E_{sample}$ depends on the regressors $\boldsymbol{X}$ and we want the effect size $E$ to be the same for the same data generating process. It is difficult to solve the effect size $E$ analytically, so we estimate it with the average of $E_{sample}$ over a very large sample size.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. These include the distribution of the regressor and the type of the simulation model (modify this according to the result).

>>>>

@@

```{=tex}
\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
\end{equation}
```
\noindent where $\Lambda(.)$ is the standard logistic function given as $\Lambda(z) = exp(z)/(1+exp(z))$. The effect size $E$ is the only regressor and the intercept is fixed to $log(0.05/0.95)$ so that $\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05$.

@@

\noindent where $M$ is a given model that generates the data, $diag(.)$ is the diagonal matrix constructed from the diagonal elements of a matrix,  $\boldsymbol{X}_a = (\boldsymbol{1},  \boldsymbol{X})$ is the matrix of regressors including the intercept used in the regression equation, $\sigma^2\boldsymbol{I}$ is the assumed variance of the error term when $H_0$ is true, $\boldsymbol{V}$ is the actual variance of the error term, and $\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z$ is the expected values of the residuals with $\boldsymbol{Z}$ be any higher order terms of $\boldsymbol{X}$ leave out by the regression equation and $\boldsymbol{\beta}_z$ be the corresponding coefficients.

A conditional expectation is taken because $E$ is designed to be invariant to different realizations of $\boldsymbol{X}$. However, solving this conditional expectation analytically is difficult as $\boldsymbol{X}$ can follow almost any distribution. We make an assumption that $\boldsymbol{X}$ follows a particular multivariate normal distribution, and repeatedly draw samples from the given model $M$. The estimate of $E$ is obtained by taking the average of the $E_{sample}$ calculated for each sample.

To study various factors contributing to the power of the visual test, the same logistic regression model is fit on different subsets of the collated data grouped by levels of factors. These include the distribution of the regressor, type of the simulation model and the shape of the residual departures.

### 2023-02-15 paper_comparison/paper_comparison/paper_comparison.tex
mention the regressor dist is assumed to be the same

(use uniform effect size? normal effect size? -\textgreater{} other
dist?)

(average out dist for effect size calc?)

@@

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size is the only
regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).

@@

\noindent where \(M\) is a given model that generates the data,
\(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of a matrix,
\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
regressors including the intercept used in the regression equation,
\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
error term, and
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of the residuals with \(\boldsymbol{Z}\) be any
variables leave out by the regression equation and
\(\boldsymbol{\beta}_z\) be the corresponding coefficients. An
expectation is taken because \(E_{sample}\) depends on the regressors
\(X\) and we want the effect size \(E\) to be the same for the same data
generating process. It is difficult to solve the effect size \(E\)
analytically, so in the analysis, we estimate it with the average of
\(E_{sample}\) over a very large sample size.

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the regressor and the type of the simulation model
(modify this according to the result).

@@

To compare the power of visual test and conventional test, we pick RESET
test (\texttt{resettest}) and Breusch-Pagan test (\texttt{bptest}) from
the R package \texttt{lmtest}, and Shapiro-Wilk test
(\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of regressors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary regressors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.

>>>>

@@

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).

@@

\noindent where \(M\) is a given model that generates the data,
\(diag(.)\) is the diagonal matrix constructed from the diagonal
elements of a matrix,
\(\boldsymbol{X}_a = (\boldsymbol{1}, \boldsymbol{X})\) is the matrix of
regressors including the intercept used in the regression equation,
\(\sigma^2\boldsymbol{I}\) is the assumed variance of the error term
when \(H_0\) is true, \(\boldsymbol{V}\) is the actual variance of the
error term, and
\(\boldsymbol{\mu}_z = \boldsymbol{R}_a\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of the residuals with \(\boldsymbol{Z}\) be any
higher order terms of \(\boldsymbol{X}\) leave out by the regression
equation and \(\boldsymbol{\beta}_z\) be the corresponding coefficients.

A conditional expectation is taken because \(E\) is designed to be
invariant to different realizations of \(\boldsymbol{X}\). However,
solving this conditional expectation analytically is difficult as
\(\boldsymbol{X}\) can follow almost any distribution. We make an
assumption that \(\boldsymbol{X}\) follows a particular multivariate
normal distribution, and repeatedly draw samples from the given model
\(M\). The estimate of \(E\) is obtained by taking the average of the
\(E_{sample}\) calculated for each sample.

To study various factors contributing to the power of the visual test,
the same logistic regression model is fit on different subsets of the
collated data grouped by levels of factors. These include the
distribution of the regressor, type of the simulation model and the
shape of the residual departures.

@@

To compare the power of visual tests and conventional tests, we simulate
200000 samples evenly from the models used in the human subject
experiments, and record the conventional \(p\)-values produced from
every sample. We pick RESET test (\texttt{resettest}) and Breusch-Pagan
test (\texttt{bptest}) from the R package \texttt{lmtest}, and
Shapiro-Wilk test (\texttt{shapiro.test}) from the built-in R package
\texttt{stats}. Among them, RESET test is the only exact and appropriate
test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk
test are approximate and inappropriate tests. Their estimated power is
shown in Figure \ref{fig:polypower}. To set up the RESET test, we
include different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of regressors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary regressors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.

To get an accurate estimate of the power of conventional tests, we
simulate 200000 samples evenly from the models used in the human subject
experiments, and record the \(p\)-values produced by every sample.
